---
title: "RMC Workshop Exercise: Week 1"
subtitle: "Tidyverse and RMarkdown Practice"
author: "YOUR NAME HERE"
output:
  prettydoc::html_pretty:
    theme: architect
    toc: true
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: inline
  wrap: false
---


# Road Injuries of Motorcyclists in Colombia

```{r include = FALSE}
# Set chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
# Load packages
if(!require(pacman)) install.packages("pacman")
"WRITE_YOUR_CODE_HERE"
```


## Police records of fatal motor accidents in Medellín


```{r}
# Data import with read_csv()
col_accidents <- "WRITE_YOUR_CODE_HERE"
```


```{r}
# Clean names
col_accidents_clean <- col_accidents %>% 
  "WRITE_YOUR_CODE_HERE"
```


```{r}
# rename() and mutate()
col_accidents_clean2 <- col_accidents_clean %>%
  "WRITE_YOUR_CODE_HERE"
```


The dataset analysed in this report provides information about `r nrow(col_accidents_clean2)` deaths caused by motorcycle accidents in Medellín, a Colombian city, from 2012 to 2015. The data is sourced from medical and police certificates and can be accessed [here](https://zenodo.org/record/4836304).

# Age and sex of motorcycle deaths

```{r}
# Create subset
col_accidents_subset <- "WRITE_YOUR_CODE_HERE"
```


```{r}
# Summarize data
age_sex_table <- "WRITE_YOUR_CODE_HERE"
```

```{r}
# Visualize summary data with ggplot() and geom_col()
"WRITE_YOUR_CODE_HERE"
```




```{r}
# Export summary data frame as a CSV to your "outputs" folder, using write_csv()
"WRITE_YOUR_CODE_HERE"
```



```{r}
# Save your ggplot to the "outputs" folder as an image file, using ggsave()
"WRITE_YOUR_CODE_HERE"
```





